Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-02-10 trunk/constantRate/ConstCompFuzzExt.tex
cleaning up claim d.3

Our first construction requires high individual entropy of a constant fraction of symbols, but permits symbols to be dependent.  Our second construction requires  a constant fraction of symbols to have a constant amount of entropy conditioned on prior symbols. The constructions can be implemented efficiently based on number-theoretic assumptions or assumptions on cryptographic hash functions.
\bnote{more about what we do}
\end{abstract}

@@

\paragraph{Limitations of Entropy-Based Approaches}
Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess some $w'$ within distance $t$ of $w$ they will be able to easily distinguish $r$ from uniform by running $\rep$. If $t$ is larger, then $\rep$ tolerates more $w'$ values, so the adversary's job is easier. In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then the \emph{entire distribution} of $w$ may have min-entropy $m$ and be contained in a single ball.  Then the adversary can always produce the key by running $\rep$ with the center of the ball as $w'$.\bnote{was told this wasn't clear enough}

More generally, let $B_t$ denote the number of points in a ball of radius $t$. For any $m$ and $t$, there is a distribution of min-entropy $m$ such that the adversary can guess a correct $w'$ with probability $1/\lceil 2^m/B_t) \rceil\approx B_t 2^{-m}$: the distribution consists of the uniform distribution over all points in several non-overlapping balls of radius $t$ (the metric space must be large enough for these balls not to intersect). We thus call $m-\log B_t$ the \emph{minimum usable} entropy, denoted by $\Huse$. The previous paragraph shows that  no fuzzy extractor can handle all distributions of a given min-entropy $m$ if  $\Huse\le 0$.

@@

We now consider the number of possible $0\rightarrow 1$ bit flips in $c$.  A $0\rightarrow 1$ flip on index $j$ occurs when two conditions are met $w_j\neq w_j'$ and $c_j = 0$.  Since the first conditioned is only fulfilled at most $t$ times, we have the following lemma:

@@

\subsection{Discussion of \consref{cons:first construction}}
\label{sec:discussion}
To show that $\Huse$ can be negative for \consref{cons:first construction}, we first calculate the size of the Hamming ball.  We allow $t$ errors with an alphabet $\mathcal{Z}$ of size $2^n$.  This means that
\[
\log |B_t| = \log \sum_{i=0}^t {\ell \choose i} (|\mathcal{Z}|-1)^i> \log {\ell \choose t} (|\mathcal{Z}|-1)^t =\Theta(t\log |\mathcal{Z}|) + \log {\ell\choose t}
\]

@@

\begin{definition}
\label{def:partial source}
A distribution $W = W_1,..., W_\gamma$ is an $(\alpha, \beta)$-partial block source if there exists a set of indices $J$ where $|J| \geq \gamma - \beta$ such that the following holds:
\[
\forall j\in J, \forall w_1,..., w_\gamma \in W_1,..., W_\gamma, \Hoo(W_j | W_1 = w_1,..., W_{j-1}=w_{j-1}) \geq \alpha.
\]
\end{definition}

\defref{def:partial source} is a weakening of block sources~(introduced by Chor and Goldreich~\cite{DBLP:journals/siamcomp/ChorG88}), as only some blocks are required to have entropy conditioned on the past.  The choice of conditioning on the past is arbitrary: a more general sufficient condition is that there exists some ordering of indices where most items have entropy conditioned on all previous items in this ordering~(for example, a ``partial'' reverse block source~\cite{vadhan2003constructing}).
%Note a partial block source requires worst case entropy while an unguessable block distribution only requires average conditional entropy. \lnote{this comparison is unclear to me}  However, we achieve security for significantly lower entropy levels than for unguessable block distributions.

Let $\sample_{\gamma, \eta}(\cdot)$ be an algorithm that  outputs a random subset of $\{1,..., \gamma\}$ of size $\eta$ given let $r_{sam}$ bits of randomness.

@@

\subsection{Discussion of \consref{cons:sampling}}
\lnote{I need to check this}
We now show \consref{cons:sampling} can work for partial block sources when $\Huse\le 0$.  The required entropy of partial block source is $\alpha (\gamma-\beta ) = \Theta(\gamma)$.  We are able to correct $O(\gamma/\eta)$ errors.
This yields:
\begin{align*}
\Huse &= \Hoo(W) -\log |B_t| \\
&< \Theta(\gamma)- t \log |\mathcal{Z}|\\
&= \Theta(\gamma) - \Theta(\gamma/\eta) \log |\mathcal{Z}|
\end{align*}
That is, \consref{cons:sampling} achieves $\Huse\le 0$ when the starting alphabet is super polynomial~(noting that for super polynomial size $\mathcal{Z}$ we can set $\eta$ to be super logarithmic and $o(\log |\mathcal{Z}|)$).  We note that for polynomial-size alphabets, \consref{cons:sampling} will still work as long as we use a code that corrects Hamming errors in both directions (with a polynomial size alphabet the probability of $1\rightarrow 0$ bits flips is noticeable); however, for a polynomial-size alphabet, $\Huse>0$.
%However, in this setting $\Huse =\Omega(\gamma)$ and known information-theoretic fuzzy extractors provide superior performance.
\bibliographystyle{alpha}
\bibliography{crypto}

@@

\section{Proof of \lemref{lem:cond and cext}}
\label{sec:cond and cext}
\lnote{I need to check this}
\begin{proof}
It suffices to show if there is some distinguisher $D'$ of size $s'$ where
\[\delta^{D'}((\cext(X; U_d), U_d, SS), (U_\kappa, U_d, SS))>\epsilon_{cond}+ \epsilon_{ext}\]
 then there is an distinguisher $D$ of size $s_{cond}$ where for all $Y$ where $\Hav(Y|SS) \geq \tilde{m}$ such that
 \[
 \delta^{D}((X, SS), (Y, SS))\geq \epsilon_{cond}.
 \]
Let $D'$ be such a distinguisher.  That is,
\[
\delta^{D'}(\cext(X, U_d)\times U_d \times SS, U_\kappa\times U_d\times SS)> \epsilon_{ext}+\epsilon_{cond}.
\]
Then define $D$ as follows.  On input $(y, ss)$ sample $seed\leftarrow U_d$, compute $r\leftarrow \cext(y; seed)$ and output $D(r, seed, ss)$.  Note that $|D| \approx s' + |\cext| +d= s_{cond}$.  Then we have the following:
\begin{align*}
\delta^{D}((X, SS), (Y, SS))&= \delta^{D'}((\cext(X, U_d), U_d, SS), \cext(Y, U_d), U_d, SS)\\
&\geq \delta^{D'}((\cext(X, U_d), U_d, SS), (U_\kappa\times U_d \times SS)) - \delta^{D'}((\cext(Y, U_d), U_d, SS), (U_\kappa\times U_d \times SS))\\
&>\epsilon_{cond}+\epsilon_{ext}- \epsilon_{ext} = \epsilon_{cond}.
\end{align*}
Where the last line follows by noting that $D'$ is of size at most $s_{ext}$.  Thus $D$ distinguishes $X$ from all $Y$ with sufficient conditional min-entropy.  This is a contradiction.
\end{proof}

@@

\bnote{better name for guess, creativity is running low.}

@@

It suffices to show that there exists an $S$ making $q$ queries such that for some $j\in J, \Hav(W_j | S^{I_{W}(\cdot, \cdot)})\le \log |\mathcal{Z}| - \log q$.  Denote by $guess_{1}, ..., guess_q$ the $q$ most likely outcomes of $W_j$~(breaking ties arbitrarily).  Note that $\sum_{i=1}^q \Pr[W_j = guess_i]\geq q/|\mathcal{Z}|$.  Suppose not, this means that there is some $guess_i$ with probability $\Pr[W_j = guess_i] < 1/|\mathcal{Z}|$.  Since there are $\mathcal{Z} - q $ remaining possible values of $W_j$ for their total probability to be at least $1-q/|\mathcal{Z}|$ at least of these values has probability at least $1/\mathcal{Z}$.  This contradicts the statement $guess_1,..., guess_q$ are the most likely values.  Consider $S$ that queries its oracle on $(j, guess_1),.., (j, guess_q)$.  Denote by $Guess$ the random variable when $W_j\in \{guess_1,.., guess_q\}$  After these queries the remaining min-entropy is at most:

@@

\Hav(W_j | S^{J_W(\cdot, \cdot)}) &=  -\log \left(\Pr[Guess=1]\times 1+ \Pr[Guess=0]\times \max_{w}\Pr[W_j = w|\not Guess =0]\right)\\
&\leq  -\log \left(\Pr[Guess=1]\times 1\right)\\

>>>>

Our first construction requires high individual entropy of a constant fraction of symbols, but permits symbols to be dependent.  Our second construction requires  a constant fraction of symbols to have a constant amount of entropy conditioned on prior symbols. The constructions can be implemented efficiently based on number-theoretic assumptions or assumptions on cryptographic hash functions.
\bnote{more about what we do}

\textbf{Keywords} Fuzzy extractors, key derivation, error-correcting codes, computational entropy, point obfuscation.
\end{abstract}

@@

\paragraph{Limitations of Entropy-Based Approaches}
\bnote{shorten this heading?}
Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess some $w'$ within distance $t$ of $w$ they will be able to easily distinguish $r$ from uniform by running $\rep$. If $t$ is larger, then $\rep$ tolerates more $w'$ values, so the adversary's job is easier. In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then the \emph{entire distribution} of $w$ may have min-entropy $m$ and be contained in a single ball.  Then the adversary can always produce the key by running $\rep$ with the center of the ball as $w'$.\bnote{was told this wasn't clear enough}

More generally, let $B_t$ denote the number of points in a ball of radius $t$. For any $m$ and $t$, there is a distribution of min-entropy $m$ such that the adversary can guess a correct $w'$ with probability $1/\lceil( 2^m/B_t) \rceil\approx B_t 2^{-m}$: the distribution consists of the uniform distribution over all points in several non-overlapping balls of radius $t$ (the metric space must be large enough for these balls not to intersect). We thus call $m-\log B_t$ the \emph{minimum usable} entropy, denoted by $\Huse$. The previous paragraph shows that  no fuzzy extractor can handle all distributions of a given min-entropy $m$ if  $\Huse\le 0$.
\bnote{is there anyway we can reasonably say what previous works support?}

@@

Since there are most $t$ locations when $w_j\neq w_j'$ there are at most $t$ $0\rightarrow 1$ bit flips in $c$:

@@

\subsection{Discussion of \consref{cons:first construction}}
\label{sec:discussion}
To show that $\Huse$ can be negative for \consref{cons:first construction}, we first calculate the size of the Hamming ball.  
\[
\log |B_t| = \log \sum_{i=0}^t {\ell \choose i} (|\mathcal{Z}|-1)^i> \log {\ell \choose t} (|\mathcal{Z}|-1)^t =\Theta(t\log |\mathcal{Z}|) + \log {\ell\choose t}
\]

@@

\begin{definition}
\label{def:partial source}
A distribution $W = W_1,..., W_\gamma$ is an $(\alpha, \beta)$-partial block source if there exists a set of indices $J$ where $|J| \geq \gamma - \beta$ such that the following holds:
\[
\forall j\in J, \forall w_1,..., w_\gamma \in W_1,..., W_\gamma, \Hoo(W_j | W_1 = w_1,..., W_{j-1}=w_{j-1}) \geq \alpha.
\]
\end{definition}
\defref{def:partial source} is a weakening of block sources~(introduced by Chor and Goldreich~\cite{DBLP:journals/siamcomp/ChorG88}), as only some blocks are required to have entropy conditioned on the past.  The choice of conditioning on the past is arbitrary: a more general sufficient condition is that there exists some ordering of indices where most items have entropy conditioned on all previous items in this ordering~(for example, a ``partial'' reverse block source~\cite{vadhan2003constructing}).
Let $\sample_{\gamma, \eta}(\cdot)$ be an algorithm that  outputs a random subset of $\{1,..., \gamma\}$ of size $\eta$ given let $r_{sam}$ bits of randomness.

@@

\subsection{Discussion of \consref{cons:sampling}}
\lnote{I need to check this}
We now show \consref{cons:sampling} can work for partial block sources when $\Huse\le 0$.  The required entropy of partial block source is $\alpha (\gamma-\beta ) = \Theta(\gamma)$.  We are able to correct $O(\gamma/\eta)$ errors.
This yields:
\[
\Huse = \Hoo(W) -\log |B_t| < \Theta(\gamma)- t \log |\mathcal{Z}|= \Theta(\gamma) - \Theta(\gamma/\eta) \log |\mathcal{Z}|
\]
That is, \consref{cons:sampling} achieves $\Huse\le 0$ when the starting alphabet is super polynomial~(noting that for super polynomial size $\mathcal{Z}$ we can set $\eta$ to be super logarithmic and $o(\log |\mathcal{Z}|)$).  We note that for polynomial-size alphabets, \consref{cons:sampling} will still work as long as we use a code that corrects Hamming errors in both directions (with a polynomial size alphabet the probability of $1\rightarrow 0$ bits flips is noticeable); however, for a polynomial-size alphabet, $\Huse>0$.
%However, in this setting $\Huse =\Omega(\gamma)$ and known information-theoretic fuzzy extractors provide superior performance.
\bibliographystyle{alpha}
\bibliography{crypto}

@@

\section{Proof of \lemref{lem:cond and cext}}
\label{sec:cond and cext}
\lnote{I need to check this}
\begin{proof}
It suffices to show if there is some distinguisher $D'$ of size $s'$ where
\[\delta^{D'}((\cext(X; U_d), U_d, SS), (U_\kappa, U_d, SS))>\epsilon_{cond}+ \epsilon_{ext}\]
 then there is an distinguisher $D$ of size $s_{cond}$ where for all $Y$ where $\Hav(Y|SS) \geq \tilde{m}$ such that
 \[
 \delta^{D}((X, SS), (Y, SS))\geq \epsilon_{cond}.
 \]
Let $D'$ be such a distinguisher.  That is,
\[
\delta^{D'}(\cext(X, U_d)\times U_d \times SS, U_\kappa\times U_d\times SS)> \epsilon_{ext}+\epsilon_{cond}.
\]
Then define $D$ as follows.  On input $(y, ss)$ sample $seed\leftarrow U_d$, compute $r\leftarrow \cext(y; seed)$ and output $D(r, seed, ss)$.  Note that $|D| \approx s' + |\cext| +d= s_{cond}$.  Then we have the following:
\begin{align*}
\delta^{D}((X, SS), (Y, SS))&= \delta^{D'}((\cext(X, U_d), U_d, SS), \cext(Y, U_d), U_d, SS)\\
&\geq \delta^{D'}((\cext(X, U_d), U_d, SS), (U_\kappa\times U_d \times SS)) \\
&- \delta^{D'}((U_\kappa\times U_d \times SS), (\cext(Y, U_d), U_d, SS))\\
&>\epsilon_{cond}+\epsilon_{ext}- \epsilon_{ext} = \epsilon_{cond}.
\end{align*}
Where the last line follows by noting that $D'$ is of size at most $s_{ext}$.  Thus $D$ distinguishes $X$ from all $Y$ with sufficient conditional min-entropy.  This is a contradiction.
\end{proof}

@@

@@

It suffices to show that there exists an $S$ making $q$ queries such that for some $j\in J, \Hav(W_j | S^{I_{W}(\cdot, \cdot)})\le \log |\mathcal{Z}| - \log q$.  Denote by $w_{j,1}, ..., w_{j,q}$ the $q$ most likely outcomes of $W_j$~(breaking ties arbitrarily).  Then $\sum_{i=1}^q \Pr[W_j = w_{j,i}]\geq q/|\mathcal{Z}|$.  Suppose not, this means that there is some $w_{j,i}$ with probability $\Pr[W_j = w_{j,i}] < 1/|\mathcal{Z}|$.  Since there are $\mathcal{Z} - q $ remaining possible values of $W_j$ for their total probability to be at least $1-q/|\mathcal{Z}|$ at least of these values has probability at least $1/\mathcal{Z}$.  This contradicts the statement $w_{j,1},..., w_{j,q}$ are the most likely values.  Consider $S$ that queries its oracle on $(j, w_{j,1}),.., (j, w_{j,q})$.  Denote by $Bad$ the random variable when $W_j\in \{w_{j,1},.., w_{j,q}\}$  After these queries the remaining min-entropy is at most:

@@

\Hav(W_j | S^{J_W(\cdot, \cdot)}) &=  -\log \left(\Pr[Bad=1]\times 1+ \Pr[Bad=0]\times \max_{w}\Pr[W_j = w| Bad =0]\right)\\
&\leq  -\log \left(\Pr[Bad=1]\times 1\right)\\
